---
title: "Homework w43"
author: "Laura Bock Paulsen"
date: "10/27/2022"
output: html_document
---

```{r setup}
pacman::p_load(dplyr, gapminder)
```

```{r}
df <-  gapminder
```

### Exercise 1: Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. You can use the population and GDPpercapita columns for it. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.
```{r}
calculate_GDP <- function(dat){
    if(any(names(dat) == c('pop', 'gdpPercap'))){
      dat$gdp = dat$pop * dat$gdpPercap
    }
    else{stop("dat must include columns'pop' and 'gdpPercap' for this computation to continue")
    }
  dat
 }
```

```{r}
df %>% filter(country == 'Denmark' & year %in% c(1967, 1977, 1987, 1997, 2007, 2017)) %>% calculate_GDP() %>% select(year, pop, gdpPercap, gdp)
```

Above the gdp calculated for Denmark in the wanted years can be seen. Do however note that the year 2017 is not included here, since it was not present in the data. 

### Exercise 2: Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and prints out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70. (Hint: remember the grepl function, and review the Control Flow tutorial)
```{r}
unique_countries = unique(df$country)

for(i in 1:length(unique_countries)){ # loops over each country
  
  if(substr(unique_countries[1], 1, 1) == 'B'){ # Checks if the first letter is  a B
    b = 'Starts with B. '             
  }
  else{
    b = 'Does not start with B. '
  }
  
  # calculating the mean life expectancy
  tmp_data = df %>% filter(country == unique_countries[i]) # temporary data with just one countries data
  mean_life = round(mean(tmp_data$lifeExp), 2)
  
  if(mean_life < 50){
    mean_life_info = '< 50'
  }
  if(mean_life>50 & mean_life<70){
    mean_life_info = 'between 50 and 70'
  }
  if(mean_life>70){
    mean_life_info = '> 70'
  }

  
  # The output for each loop
  print(paste0(unique_countries[i],': ', b, 'The mean life expectancy is: ', mean_life, ', which means that it is ', mean_life_info))
}
```


## Exercise 3:
Challenge/Optional: Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘M’ and graphs life expectancy against time (using plot() function) as a line graph if the mean life expectancy is under 50 years.
Hint: If you are struggling with the gapminder tibble format, consider converting it into a dataframe, either by downloading it from the internet and loading it via read.csv (not read_csv), and/or using as.data.frame() conversion function and then appropriately subsetting.