---
title: "Starting with R - Homework"
author: "Laura Bock Paulsen"
date: "8/30/2022"
output: html_document
---

### 1) Use R to figure out how many elements in the vector below are greater than 2 and then tell me what their sum (of the larger than 2 elements) is.

```{r}
rooms <- c(1, 2, 4, 5, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, 9, 3, NA)

rooms_noNA <- na.omit(rooms)
```

```{r}
sum(rooms_noNA > 2)
```

Explanation of why the above works:
1. The NA's are omitted
2. Then we check for each element of the vector whether the number is larger than zero. This returns a new vector containing either TRUE or FALSE for each original element
```{r}
rooms_noNA > 2
```

3. As FALSE and TRUE also can be seen as 0 and 1's we find the number of items greater than 2 by taking the sum of the vector. 


Then their sum is calculated:
```{r}
sum(rooms_noNA[rooms_noNA > 2])
```



### 2) What type of data is in the 'rooms' vector?
```{r}
class(rooms)
```


### 3) Submit the following image to Github: Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing
#### a) the line of code you used to create the object,
```{r, warning=FALSE}
# i prefer using the p_load function from the package pacman as it both opens the package as well as automatically installing the package if you don't have it yet.
pacman::p_load('tidyverse')

# downloading file
download.file("../https://ndownloader.figshare.com/files/11492171",
             "data/SAFI_clean.csv", mode = "wb")

# reading the file into R
interviews <- read.csv('../data/SAFI_clean.csv')
```



#### b) the 'interviews' object in the Environment, and
#### c) the file structure of your R project in the bottom right "Files" pane.
### d) Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS) or equivalent. Place here the URL leading to the screenshot in your repository.

4) Challenge: Attached to this assignment you will find Danish monarchs dataset that might need to be cleaned up a bit. Load the provided 'kings.csv" dataset into R as a tibble. Calculate the mean() and median() duration of rule over time and find the three mondarchs ruling the longest. How many days did they rule (accounting for transition year?)

```{r, warning=FALSE}
# Using read.csv2 to indicate that the data is semicolon seperated (read.csv works for commaseperated files)
setwd('..')
kings <- read.csv2("data/kings.csv")
```

Calculating the duration which will be used for calculating the mean and median
```{r}
duration <- kings$End_date - kings$Start_date
```

```{r}
mean(duration, na.rm = T) # using na.rm to exclude NA's from the calculation
```

```{r}
median(duration, na.rm = T)
```

```{r}
max(duration, na.rm = T)
```

